### 2014-06-04 writing/S'14 paper/intro.tex
First minor tweaks from the review session.

In our experiments, we use this pre-specified logical grammar to
generate controlled data sets encoding the semantic relationships
between pairs of expressions and evaluating whether each of two
classes of neural model --- plain RNNs and recursive neural tensor
networks (RNTNs, \cite{socher2013acl1}) --- can learn those
relationships correctly. In our first experiment
(Section~\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (Section~\ref{sec:recursion}) extends this
evaluation to cover relations between complex recursive structures
like $(a \plor b)$ and $\plneg(\plneg a \pland \plneg b)$, and our
third experiment (Section~\ref{sec:quantifiers}) involves relations
between quantified statements like \ii{every reptile walks} and \ii{every
turtle moves}. We find that the plain RNN achieves only mixed results
in all three experiments, whereas the stronger RNTN models generalized
well in every case, suggesting that it has in fact learned, or at
least learned to simulate, our target logical
concepts. % TODO: Update with current results

>>>>

In our experiments, we use this pre-specified logical grammar to
generate controlled data sets encoding the semantic relationships
between pairs of expressions and evaluate whether each of two
classes of neural model --- plain RNNs and recursive neural tensor
networks (RNTNs, \cite{socher2013acl1}) --- can learn those
relationships correctly. In our first experiment
(Section~\ref{sec:join}), we evaluate the ability of these models to
learn the core relational algebra of natural logic from data. Our
second experiment (Section~\ref{sec:recursion}) extends this
evaluation to cover relations between complex recursive structures
like $(a \plor b)$ and $\plneg(\plneg a \pland \plneg b)$, and our
third experiment (Section~\ref{sec:quantifiers}) involves relations
between quantified statements like \ii{every reptile walks} and \ii{every
turtle moves}. We find that the plain RNN achieves only mixed results
in all three experiments, whereas the stronger RNTN models generalized
well in every case, suggesting that it has in fact learned, or at
least learned to simulate, our target logical
concepts. % TODO: Update with current results

### 2014-06-04 writing/S'14 paper/methods.tex
First minor tweaks from the review session.

We study neural models that adhere to the linguistic \ii{principle of
  compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their constituent parts
via specific combination functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are vectors of length $N$. The
combination function concatenates pairs of them to form vectors of
length $2N$ and maps those larger vectors back into length $N$
vectors, which can then be merged again to create more complex
phrases. Once the entire sentence-level representation has been
derived, it fed to a classifier trained in a supervised manner.

@@

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and the RNTN layer function
proposed in \citet{chen2013learning} \eqref{rntn}. A sigmoid
nonlinearity (element-wise $\tanh$) is applied to the output of either
layer function, following \cite{socher2013acl1}.
%
\begin{gather} \label{rnn}
y_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
\label{rntn}
y_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{A}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
\end{gather} % TODO: Explain third order tensor parameter?
%
Here, the $l$ and $r$ superscripts indicate the vector representations
for the left and right children of the local structure interpreted as
$y$. The RNN concatenates them, multiplies them by an $N \times 2N$
matrix of learned weights, and applies the element-wise non-linearity
to the resulting vector (plus a bias term). The RNTN has the same
basic structure, but with the addition of a tensor $\mathbf{A}$,
dimension $N \times N \times N$, modeling interactions between the
child vectors.

>>>>

We study neural models that adhere to the linguistic \ii{principle of
  compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their constituent parts
via specific combination functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are vectors of length $N$. The
combination function concatenates pairs of them to form vectors of
length $2N$ and maps those larger vectors back into length $N$
vectors, which can then be merged again to create more complex
phrases. Once the entire sentence-level representation has been
derived, it is fed to a classifier trained in a supervised manner.

@@

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and the RNTN layer function
proposed in \citet{chen2013learning} \eqref{rntn}. A sigmoid
nonlinearity (element-wise $\tanh$) is applied to the output of either
layer function, following \cite{socher2013acl1}.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
\label{rntn}
\vec{y}_{\textit{RNTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots N]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
\end{gather} % TODO: Explain third order tensor parameter?
%
Here, the $\vec{x}^{(l)}$ and $\vec{x}^{(l)}$ are the column vector representations
for the left and right children of the node, and $\vec{y}$ is the node's output.
The RNN concatenates them, multiplies them by an $N \times 2N$
matrix of learned weights, and applies the element-wise non-linearity
to the resulting vector (plus a bias term). The RNTN has the same
basic structure, but with the addition of a learned third-order tensor $\mathbf{T}$,
dimension $N \times N \times N$, modeling multiplicative interactions between the
child vectors. Both models include a learned bias vector, represented $\vec{b}$.

### 2014-06-04 writing/S'14 paper/quantifiers.tex
First minor tweaks from the review session.

Even to the extent that RNTN models can handle functional meanings in
the form of the operators of propositional logic, this is not a
guarantee that these models can handle functional meanings of the form
seen in natural language. As a first step towards investigating this
issue, we attempt to directly measure the degree to which RNNs are
able to develop suitable representations for natural language
quantifiers like \ii{some} and \ii{all}. Quantification is far from
the only place in natural language where complex functional meanings
are found, but it is a natural starting point, since it can be tested
in sentences whose structures are otherwise quite simple, and since it
has formed a standard case study in prior formal work on natural
language inference.

@@

\mynote{Discuss results of this experiment}

While our primary interest is in discovering the extent to which our
models' can learn to encode the logic given an arbitrary amount of
data, we are also interested in the degree to which they can infer a
correct representation from the logic from more constrained training
data. To this end, we segment the sentence pairs according to which
quantifiers appear in each pair, and then hold out one such pair for
testing. We hypothesize that a model that can efficiently learn to
represent a logic should be able to construct an accurate
representation of each held out quantifier from the way that it
interacts with the other nine quantifiers which are not held
out. Since running this experiment requires choosing a pair of
quantifiers to hold out before training, the resource demands of
training prevent us from testing each of the 55 possible possible
pairs of quantifiers, and we choose only four pairs to test on.  Three
of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
\ii{all}/\ii{some}) were chosen because they determine the most
diverse set of logical relationship in our data, thereby providing the
most complex training and testing scenario possible. The fourth is a
self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
handles equality.

@@

\begin{table}[htp]
  \centering
  \begin{tabular}{ l rrr }
    \toprule
    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
    \midrule
    \textsc{all split}	& 35.4 &	60.4&	\textbf{94.4}
    \\[1ex]    
    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	\textbf{97.1} \\
    \textsc{pair not-all/not-most}	&0 &	   65.6  &	\textbf{79.2} \\
    \textsc{pair all/some}	& 0& 60.7  &	\textbf{79.5} \\
    \textsc{pair no/no}	& 30.8 &	65.8 &	\textbf{100} \\
    \bottomrule
  \end{tabular}
  \caption{Quantifier experiment performance. 
    The most frequent class baseline accuracy reflects the most frequent class in the training data, $\natind$.}
  \label{resultstable}
\end{table}

\mynote{
The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
% TODO: Update dimensionality
(Note: Models still running. Results might improve and this para is
likely to change.) The perfect performance by the RNTN on the
\textsc{pair no/no} experiment and its strong performance on
\textsc{pair two/less-than-two} suggests that it is at least plausible
that this model is able to handle quantifiers correctly given
sufficient training data, but the sub 90\% results on two of the
training settings suggest that the model may not be able to
generalize from data this impoverished in general. However, the fact
that both models perform far above baseline is promising, and the
question of how much data is necessary to accurately capture
quantifier behavior in a na\"ive model remains open.

>>>>

Even to the extent that RNTN models can handle functional meanings in
the form of the operators of propositional logic, this is not a
guarantee that these models can handle functional meanings of the form
seen in natural language. As a first step towards investigating this
issue, we attempt to directly measure the degree to which RNNs are
able to develop suitable representations for the semantics of natural language
quantifiers like \ii{some} and \ii{all}. Quantification is far from
the only place in natural language where complex functional meanings
are found, but it is a natural starting point, since it can be tested
in sentences whose structures are otherwise quite simple, and since it
has formed a standard case study in prior formal work on natural
language inference.

@@

While our primary interest is in discovering the extent to which our
models can learn to encode the logic given an arbitrary amount of
data, we are also interested in the degree to which they can infer a
correct representation for the logic from more constrained training
data. To this end, we segment the sentence pairs according to which
quantifiers appear in each pair, and then hold out one such pair for
testing. We hypothesize that a model that can efficiently learn to
represent a logic should be able to construct an accurate
representation of each held out quantifier from the way that it
interacts with the other nine quantifiers which are not held
out. Since running this experiment requires choosing a pair of
quantifiers to hold out before training, the resource demands of
training prevent us from testing each of the 55 possible possible
pairs of quantifiers, and we choose only four pairs to test on.  Three
of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
\ii{all}/\ii{some}) were chosen because they provide allow for the most
different class labels at in the training data. The fourth is a
self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
handles equality.

@@

\begin{table}[htp]
  \centering
  \begin{tabular}{ l rrr }
    \toprule
    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
    \midrule
    \textsc{all split}	& 35.4 &	60.4&	\textbf{94.4}
    \\[1ex]    
    \textsc{pair two/less-than-two}	& 59.8 &	76.4 &	\textbf{97.1} \\
    \textsc{pair not-all/not-most}	&0 &	   65.6  &	\textbf{79.2} \\
    \textsc{pair all/some}	& 0& 60.7  &	\textbf{79.5} \\
    \textsc{pair no/no}	& 30.8 &	65.8 &	\textbf{100} \\
    \bottomrule
  \end{tabular}
  \caption{Quantifier experiment performance. }
  \label{resultstable}
\end{table}

The results for these experiments are shown at the top of Figure \ref{resultstable}. We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.

\mynote{
The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
% TODO: Update dimensionality
(Note: Models still running. Results might improve and this para is
likely to change.)

The perfect performance by the RNTN on the
\textsc{pair no/no} experiment and its strong performance on
\textsc{pair two/less-than-two} suggests that it is at least plausible
that this model is able to handle quantifiers correctly given
sufficient training data, but the sub 90\% results on two of the
training settings suggest that the model may not be able to
generalize from data this impoverished in general. However, the fact
that both models perform far above baseline is promising, and the
question of how much data is necessary to accurately capture
quantifier behavior in a na\"ive model remains open.

### 2014-06-04 writing/S'14 paper/recursion.tex
First minor tweaks from the review session.

Recursive structure is an prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in testing.

@@

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing depth (and with the increasing complexity of the expressions),
 resulting in the gradually dropping performance. This is not
necessarily a significant flaw in the model, since it remains possible
that a less lossy approximation of the same logic would do no better
on the training objective, since that objective only considers short
strings and includes a regularization term. Whether a more accurate
encoding could be learned from a larger set of training data remains a
quention for future work. In contrast, the RNN model did not appear to
be able to learn a correct approximation of the logic.

>>>>

Recursive structure is a prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in testing.

@@

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing depth (and with the increasing complexity of the expressions),
 resulting in the gradually dropping performance. This is not
necessarily a significant flaw in the model, since it remains possible
that a less lossy approximation of the same logic would do no better
on the training objective, since that objective only considers short
strings and includes a regularization term. Whether a more accurate
encoding could be learned from a larger set of training data remains a
question for future work. In contrast, the RNN model did not appear to
be able to learn a correct approximation of the logic.
